---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# envi

<!-- badges: start -->
[![Build Status](https://travis-ci.org/kaneplusplus/envi.svg?branch=master)](https://travis-ci.org/kaneplusplus/envi)
[![codecov](https://codecov.io/gh/kaneplusplus/envi/branch/master/graph/badge.svg)](https://codecov.io/gh/kaneplusplus/envi)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The goal of the `envi` package is manage mulitple R environments (collections of package). This means providing functionality to retrieve remote environments, create new environments, and switch between environments. The `envi` package integrates functionality provided by Kevin Ushey's [`renv` package](https://rstudio.github.io/renv/articles/renv.html), which provides individual environment fuctionality, and other packages including [`git2r`](https://docs.ropensci.org/git2r/) and [`piggyback`](https://github.com/ropensci/piggyback), which provides source control and access to remote environments.

To draw an analogy from the Python community, just as `renv` is similar to [virtual environments](https://docs.python.org/3/tutorial/venv.html), `envi` is similar to [conda](https://conda.io/en/latest/). However, the scope is more limited in the following ways.

1. Conda is a general application independent of a programming language or environment and configuration is generally performed from the terminal. The `envi` package is used within the R programming environment. This has the advantage of making it more accessible to R users, who may not be have the system administration experience needed to tune installations or troubleshoot problems with paths, for example. On the other hand, `envi` environment configurations and the R environments they manage can easily be modified by more savvy users. 
2. Conda functions as both a package manager, allowing you install programming languages environments, as well as a manager for programming-language specific libraries. While `envi` package allows you to configure a Python environment through `renv`, it is intended to create useful R environments. This limitation in scope makes the management of package dependencies, on of Conda's core goals, almost trivial because of the R commuities enforcement of downstream package compatibility via CRAN.

## Installation

The `envi` package is still under developement and is not available on CRAN. However, it can be installed from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("kaneplusplus/envi")
```
## Example

Suppose you'd like to create a deep learning model with the `keras` but you don't have Python installed, you do have Python installed but your're not sure how to get R to talk with Python, or you're feeling lazy. Instead of configuring your `keras` installation, you can download and manage an enironment on Github with `envi` on R version 2.6.1. You can check your version with `Sys.getenv()[['R_VERSION']]`

```{r}
library(envi)

# Get the appropriate repo for the platform.
env_image <- switch(Sys.info()[['sysname']],
                    Darwin = "keras-environmentx86_64-apple-darwin15.6.0.tar.bz2",
                    Windows = "keras-environmentx86_64-w64-mingw32.zip",
                    Linux = "keras-environmentx86_64-pc-linux-gnu.tar.bz2")

# Get the environment and call it keras-env. Note that the environment is big 
# and the following can take some time.
envi_pb_install(env_image, repo = "kaneplusplus/keras-envi", handle = "keras-env")

# See which environments we have and where they are located.
envi_list()
```

Now that we have the environment, which includes `keras` and an installation of Python, let's use it to build a simple deep learner.

```{r}
# Load the reticulate first.
library(reticulate)
library(envi)

# Activate the enviroment.
envi_activate("keras-env")

# Now we can use the environments packages and other functionality.
library(keras)

# Create a model to predict iris species.
dl_model <- keras_model_sequential() %>%
  layer_dense(units = 4) %>%
  layer_dense(units = 3, activation = "softmax")

# Compile the model.
dl_model %>% compile(
  loss = loss_categorical_crossentropy,
  optimizer = optimizer_adadelta(),
  metrics = c("accuracy"))

# Create the design matrix and dependent variables.
form <- Species ~ . -1
iris_x <- model.matrix(form, iris)
iris_y <- to_categorical(as.numeric(iris$Species) - 1)

# Fit the model.
dl_model %>% fit(
  iris_x,
  iris_y,
  batch_size = 150,
  epoch = 1000,
  validation_split = 0.1)

# Calculate the model accuracy.
dl_acc <-
  apply(predict(dl_model, iris_x), 1, which.max) == as.numeric(iris$Species)
sum(dl_acc) / length(dl_acc)

```

# Code of Conduct 
Please note that the 'envi' project is released with a
[Contributor Code of Conduct](CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.
